---
title: "Contingency_table_R"
author: "Kristina Bratkova"
format: 
  html:
    toc: true
    toc-location: left
    toc-depth: 4
    code-tools: true
    code-fold: true
    code-link: true
editor: visual
execute: 
  warning: false
  message: false
---

# Contingency_table_R

## Load libraries and data

```{r,message=FALSE}
library(here)
library(tidyverse)
# load data
# check here() file path, which should be Contingency_table_R directory
df <- read_csv(here("data","randomized_survey_data.csv"))
```

## Plot 1: Average opinion on the ban

First example shows average value of selected numerical variable for categorical variables on x and y-axis. The opinion on the motorbike ban was recoded so that "strongagree"=1, "agree"=0.75, "neutral"=0.5, "disagree"=0.25 and "strongdisagree"=0.

```{r}
# do contingency table of opinion on the ban for different age groups and occupations
df %>% group_by(occup, age) %>% 
  summarise(n=round(sum(opinion_ban1)/n(),2)) %>%
  ggplot(aes(x = occup, y = age, fill = n ))  +
  geom_tile(color = "white") +
  scale_fill_gradient2(high = "olivedrab", low = "tomato",
                       mid="lightyellow",midpoint = 0.5,
                       space = "Lab", name="count"
                       ,limits=c(0,1)) +
  # make sure x and y have the same scaling
  coord_fixed() +
  # add the values to the tiles, using row and name as coordinates 
  geom_text(aes(occup,age,label=n), color = "black", size = 4) +
theme_minimal() +
       # adjust text direction on x-axis 
       theme(axis.text.x = element_text(angle = 45, vjust = 1,hjust = 1),
        text = element_text(size=15)) +
       # remove the axis labels by making them blank
       xlab("Occupation") + ylab("Age") + 
       guides(fill=guide_legend(title="Average Opinion")) 
```

The above table shows average opinion of given tile subgroup.

### Another example of average opinion

We could examine the average opinion of subgroups of different trip frequency and purpose of the trip recorded in the survey.

```{r}
# do contingency table of opinion on the ban for trip frequency and purpose
df %>% group_by(freqmon1, age) %>% 
  summarise(n=round(sum(opinion_ban1)/n(),2)) %>%
  ggplot(aes(x = freqmon1, y = age, fill = n ))  +
  geom_tile(color = "white") +
  scale_fill_gradient2(high = "olivedrab", low = "tomato",
                       mid="lightyellow",midpoint = 0.5,
                       space = "Lab", name="count"
                       ,limits=c(0,1)) +
  # make sure x and y have the same scaling
  coord_fixed() +
  # add the values to the tiles, using row and name as coordinates 
  geom_text(aes(freqmon1,age,label=n), color = "black", size = 3) +
theme_minimal() +
       # adjust text direction on x-axis 
       theme(axis.text.x = element_text(angle = 45, vjust = 1,hjust = 1),
       text = element_text(size=15)) +
       # remove the axis labels by making them blank
       ylab("Age") +xlab("Trip frequency per month") + 
       guides(fill=guide_legend(title="Average Opinion"))

```

## Plot 2: Occupation and type of occupation demographic group counts

```{r}
# do contingency table of opinion on the ban for different occupations and types of occupation
df %>% group_by(occup, ft_pt) %>% 
  summarise(n=n()) %>%
  ggplot(aes(x = occup, y = ft_pt, fill = n ))  +
  geom_tile(color = "white") +
  scale_fill_viridis_c(alpha=0.7) +
  # scale_fill_gradient2(high = "olivedrab", low = "tomato",
  #                      mid="lightyellow",midpoint = 0.5,
  #                      space = "Lab", name="count"
  #                      ,limits=c(0,1)) +
  # make sure x and y have the same scaling
  coord_fixed() +
  # add the values to the tiles, using row and name as coordinates 
  geom_text(aes(occup,ft_pt,label=n), color = "black", size = 3) +
theme_minimal() +
       # adjust text direction on x-axis 
       theme(axis.text.x = element_text(angle = 45, vjust = 1,hjust = 1),
        text = element_text(size=15)) +
       # remove the axis labels by making them blank
       xlab("Occupation") + ylab("Type of occupation") + 
       guides(fill=guide_legend(title="Counts of respondents")) 
```

## Plot 3: Cumulative sum of values (number of motorbikes) for a given subgroup (occupation,type of occupation)

The household survey contains a question about the number of motorbikes owned ranging from 0 to 5. The below plot sums these value for each demographic group.

```{r}
# do contingency table of total number of motorbikes owned for different occupations and types of occupation
df %>% group_by(occup, ft_pt) %>% 
  summarise(n=round(sum(own_motob),1)) %>%
  ggplot(aes(x = occup, y = ft_pt, fill = n ))  +
  geom_tile(color = "white") +
  scale_fill_viridis_c(alpha=0.7) +
  # scale_fill_gradient2(high = "olivedrab", low = "tomato",
  #                      mid="lightyellow",midpoint = 0.5,
  #                      space = "Lab", name="count"
  #                      ,limits=c(0,1)) +
  # make sure x and y have the same scaling
  coord_fixed() +
  # add the values to the tiles, using row and name as coordinates 
  geom_text(aes(occup,ft_pt,label=n), color = "black", size = 3) +
theme_minimal() +
       # adjust text direction on x-axis 
       theme(axis.text.x = element_text(angle = 45, vjust = 1,hjust = 1),
        text = element_text(size=15)) +
       # remove the axis labels by making them blank
       xlab("Occupation") + ylab("Type of occupation") + 
       guides(fill=guide_legend(title="Counts of respondents")) 
```
